Cover letter for the original IJRR submission

### 2017-04-19 body/conclusion.tex
cosmetics

\subsection{Addressing $\mathcal{P}_3$}
This paper focuses on contact planning of multi-contact locomotion. While it does not address directly the continuous motion generation,
we had to prove that the plans produced by our planner can be executed by the considered robots. Problem $\mathcal{P}_3$ remains an active research topic~\cite{Carpentier2016, herzog2015trajectory}, for which the implemented a solution in this paper, described in appendix 
%~ and most existing solutions do not consider collisions with the environment during the planning, making it hard to validate constrained problems such as the polaris scenario
%~ or the table egress in \cite{DBLP:conf/iser/EscandeKMG08}.
The companion video shows the results that we obtained in simulation and on the real HRP-2 robot.
In these scenarios the robot is able to execute our plans, while respecting the constraints inherent to the problem: joint kinematics and torque limits and
collision avoidance.
%~ \adnote{Probably you should add in appendix an explanation of how you solve $P_3$ and refer to it here.}

>>>>

\subsection{Addressing $\mathcal{P}_3$}
This paper focuses on contact planning of multi-contact locomotion.It does not address directly the continuous motion generation,
. Problem $\mathcal{P}_3$ remains an active research topic~\cite{Carpentier2016, herzog2015trajectory},
for which no off-the shelf solution exists thus far. We thus implemented a dedicated solution, described in Appendix~\ref{app:optim}.
The companion video shows the results that we obtained in simulation and on the real HRP-2 robot.
In these scenarios the robot is able to execute our plans, while respecting the constraints inherent to the problem: joint kinematics and torque limits and
collision avoidance.
%~ \adnote{Probably you should add in appendix an explanation of how you solve $P_3$ and refer to it here.}

### 2017-04-19 body/intro.tex
cosmetics

Obtaining reasonable computation times is critical for a contact planner to be of any practical use. Our work thus aims at breaking the complexity of the acyclic contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the combinatorial of the contact selection problem. First we introduce a low-dimensional space, called the \textit{contact reachable} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact reachable} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling presents drawbacks and advantages discussed in previous related literature, which we summarize in the following.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

Most of the papers that followed the work of Bretl have explored alternative formulations to handle the combinatorial issue. Two main directions have been explored. \textbf{On the one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the method is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. Deits and Tedrake proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered, and equilibrium is not considered~\cite{DBLP:conf/humanoids/DeitsT14}. 
Dai et al.\cite{dai2014whole} extended the work of Posa et al.\cite{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
In addition to the limits of the current implementations, optimization-based approaches only converge locally.

>>>>

Obtaining reasonable computation times is critical for a contact planner to be of any practical use. Our work thus aims at breaking the complexity of the acyclic contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the combinatorics of the contact selection problem. First we introduce a low-dimensional space, called the \textit{contact reachable} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact reachable} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling presents drawbacks and advantages discussed in previous related literature, which we summarize in the following.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

Most of the papers that followed the work of Bretl have explored alternative formulations to handle the combinatorics. Two main directions have been explored. \textbf{On the one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the method is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. Deits and Tedrake proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered, and equilibrium is not considered~\cite{DBLP:conf/humanoids/DeitsT14}. 
Dai et al.\cite{dai2014whole} extended the work of Posa et al.\cite{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
In addition to the limits of the current implementations, optimization-based approaches only converge locally.

### 2017-04-19 body/results.tex
cosmetics

\begin{table*}
\centering
\begin{tabular}{ l | c}
  &  Path planning success rate \\
 \hline
   Stairs     	& 100\% \\
   Standing			& 68\% 	\\
   Car 			& 77\% 	\\
   Rubble 				& 97\% 	\\
   Race        & 88.0\% 	\\
 \end{tabular}
\caption{ Percentage of successful complete contact planning rates for each scenario, rounded to the first decimal.}
\label{tab:sucess_planning}
\quad
\end{table*}

\begin{table*}
\centering
\begin{tabular}{ l | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | c}
  &  Equilibrium success rate & Kinematic failure & Equilibrium failure \\
 \hline
   Steep stairs 	& 99.5\%  & 0.1\% 	& 0.4\% \\
   Standing up 		& 87.8\%  & 6.1\% 	& 6.1\% \\
   Car egress 		& 66.2\%  & 15.9\% 	& 17.9\% \\
   Rubble 			& 97.54\% & 0.16\% 	& 2.3\% \\
   Obstacle race 	& 92.4\%  & 0.15\% 	& 7.45\% \\
 \end{tabular}
\caption{Success rates obtained for the generation of static equilibrium contact configurations for each scenario, rounded to the first decimal. Column 1 indicates 
indicates the rate of contact generation that suceeeded. In the cases where the generation fails, it can be
either a kinematic issue (column 2), or because no contact configuration led to a static equilibrium configuration (column 3). Note that a failure in the contact generation
is not equivalent to a failure of the contact planning algorithm.}
\label{tab:requestpercent}
\quad
\end{table*}

@@

\begin{table*}
\centering
\begin{tabular}{ c | c | c }
 Scenario & Method  & Complete computation time \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & Hauser~\cite{Hauser06usingmotion} &  5.42 min  \\							 
							  & Mordatch et al.\cite{Mordatch:2012:DCB:2185520.2185539} & 2 to 10 min \\
							 & \textbf{Ours} + \cite{Carpentier2016}  & \textbf{$ <$ 2s} \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & Hauser~\cite{Hauser06usingmotion} &  4.08 min  \\
							 & Mordatch et al.\cite{Mordatch:2012:DCB:2185520.2185539} & 2 to 10 min \\
							 & \textbf{Ours}  & \textbf{$ <$ 2s}   \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & Hauser \cite{Hauser06usingmotion} &  10.08 min  \\
							 & Mordatch et al.\cite{Mordatch:2012:DCB:2185520.2185539} & 2 to 10 min \\
							 & \textbf{Ours}   & \textbf{$ <$ 5s}   \\
 \hline
   \multirow{2}{*}{Table (car) egress} & Bouyarmane et al.~\cite{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 3.5 hours  \\
							 & \textbf{Ours}  & \textbf{$<$ 60 s} \\

 \end{tabular}
\caption{Comparison between the computation times obtained by our method and previous ones for addressing the whole problem.}
\label{tab:compprev}
\quad
\end{table*}

>>>>

\begin{table}
\centering
\begin{tabular}{ l | c}
  &  Path planning success rate \\
 \hline
   Stairs     	& 100\% \\
   Standing			& 68\% 	\\
   Car 			& 77\% 	\\
   Rubble 				& 97\% 	\\
   Race        & 88.0\% 	\\
 \end{tabular}
\caption{ Percentage of successful complete contact planning rates for each scenario, rounded to the first decimal.}
\label{tab:sucess_planning}
\quad
\end{table}

\begin{table}
\centering
\begin{tabular}{ l | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{35pt} | >{\centering\arraybackslash}m{35pt} | c}
  &  Equilibrium success rate & Kinematic failure & Equilibrium failure \\
 \hline
   Steep stairs 	& 99.5\%  & 0.1\% 	& 0.4\% \\
   Standing up 		& 87.8\%  & 6.1\% 	& 6.1\% \\
   Car egress 		& 66.2\%  & 15.9\% 	& 17.9\% \\
   Rubble 			& 97.54\% & 0.16\% 	& 2.3\% \\
   Obstacle race 	& 92.4\%  & 0.15\% 	& 7.45\% \\
 \end{tabular}
\caption{Success rates obtained for the generation of static equilibrium contact configurations for each scenario, rounded to the first decimal. Column 1 indicates 
indicates the rate of contact generation that suceeeded. In the cases where the generation fails, it can be
either a kinematic issue (column 2), or because no contact configuration led to a static equilibrium configuration (column 3). Note that a failure in the contact generation
is not equivalent to a failure of the contact planning algorithm.}
\label{tab:requestpercent}
\quad
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{ c | c | c }
 Scenario & Method  & Computation time \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & Hauser~\cite{Hauser06usingmotion} &  5.42 min  \\							 
							  & Mordatch et al.\cite{Mordatch:2012:DCB:2185520.2185539} & 2 to 10 min \\
							 & \textbf{Ours} + \cite{Carpentier2016}  & \textbf{$ <$ 2s} \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & Hauser~\cite{Hauser06usingmotion} &  4.08 min  \\
							 & Mordatch et al.\cite{Mordatch:2012:DCB:2185520.2185539} & 2 to 10 min \\
							 & \textbf{Ours}  & \textbf{$ <$ 2s}   \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & Hauser \cite{Hauser06usingmotion} &  10.08 min  \\
							 & Mordatch et al.\cite{Mordatch:2012:DCB:2185520.2185539} & 2 to 10 min \\
							 & \textbf{Ours}   & \textbf{$ <$ 5s}   \\
 \hline
   \multirow{2}{*}{Table (car) egress} & Bouyarmane et al.~\cite{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 3.5 hours  \\
							 & \textbf{Ours}  & \textbf{$<$ 60 s} \\

 \end{tabular}
\caption{Comparison between the computation times obtained by our method and previous ones for addressing the whole problem.}
\label{tab:compprev}
\quad
\end{table}
